True versus apparent shapes of astrophysical bow shocks

### 2017-04-03 papers/Paper1/quadrics-appendix.tex
Started discussion of false negative rate, beta
Wrote a bit about effect sizes

The \(p\)-values are the probability of finding a difference between
two populations as large as (or larger than) what is observed,
\emph{given} that there is no difference in the underlying
distribution from which the two populations are drawn (that is, given
that the null hypothesis is true).  However, what we really want to
know is something else: the probability that the null hypothesis is
true, \emph{given} the observations.  That is, the probability,
\(\alpha\), of a \textit{false positive}, also known as the \textit{Type I
  error rate}.  The common mistake of conflating these two definitions
is known as the ``\(p\)-value fallacy'' \citep{Goodman:1999a}, or``the
error of the transposed conditional'', as discussed in detail by
\citet{Colquhoun:2014a}.  It is possible to derive \(\alpha\) from
\(p\) using Bayes' theorem (e.g., \citealp{Goodman:1999b}), but that
requires an estimate of the prior probability of the null hypothesis,
independent of the observations.  Alternatively, it is also possible to find a lower bound on \(\alpha\) from a frequentist approach \citep{Sellke:2001a}:
\begin{equation}
  \label{eq:type-I}
  \alpha(p) \ge \bigg[ 1 - \big(e p \ln p\big)^{-1} \bigg]^{-1}
\end{equation}

@@

>>>>

@@

The \(p\)-values are the probability of finding a difference between
two populations as large as (or larger than) what is observed,
\emph{given} that there is no difference in the underlying
distribution from which the two populations are drawn (that is, given
that the null hypothesis is true).  However, what we really want to
know is something else: the probability that the null hypothesis is
true, \emph{given} the observations.  That is, the probability,
\(\alpha\), of a \textit{false positive}, also known as the \textit{Type I
  error rate}.  The common mistake of conflating these two definitions
is known as the ``\(p\)-value fallacy'' \citep{Goodman:1999a}, or``the
error of the transposed conditional'', as discussed in detail by
\citet{Colquhoun:2014a}.  It is possible to derive \(\alpha\) from
\(p\) using Bayes' theorem (e.g., \citealp{Goodman:1999b}), but that
requires an estimate of the prior probability of the null hypothesis,
independent of the observations.  Alternatively, it is also possible
to find a lower bound on \(\alpha\) from a frequentist approach
\citep{Sellke:2001a}:
\begin{equation}
  \label{eq:type-I}
  \alpha(p) \ge \bigg[ 1 - \big(e\, p \ln p\big)^{-1} \bigg]^{-1}
  \quad \text{valid for } p < 1/e.
\end{equation}
This is the approach we adopt here, which also numerically coincides
with the Bayesian approach for the case where the prior probability of
the null hypothesis is 0.5.  The reason that this is only a lower
limit for \(\alpha\) is that if we have overwhelming a priori evidence that
the null hypothesis is true (for instance, from previous empirical
studies, or because it follows from a well-supported theory), then a
Bayesian calculation would give a much higher value of \(\alpha\) than
\eqref{eq:type-I} does.  In our case, however, we have no strong
reasons for favoring any of the null hypotheses, so it is reasonable
to assume \(\alpha\) is close to the lower limit of \eqref{eq:type-I}.

In order to choose a threshold \(p\)-value that counts as a
``significant'' result, one then needs to balance the risks of false
positives against the risks of \textit{false negatives}.  The false
negative probability, \(\beta\), also known as \textit{Type II error
  rate}, is the probability of failing to reject an untrue null
hypothesis.  That is, in the context of this paper, it is the
probability of failing to detect a real difference between two
sub-samples, or a real correlation between two variables.  The
complementary probability, \(1 - \beta\), is known as the
\textit{statistical power} or sensitivity of the test.  The value of
\(\beta\) depends on three factors:
\begin{enumerate}[1.]
\item The \textit{effect size}, which is a measure of the magnitude of
  the difference in a dependent variable between two sub-samples, or
  the degree of correlation between two continuous variables.  For the
  two sub-sample case, it is common to use a standardised mean
  difference, such as Cohen's \(d\) statistic \citep{Cohen:1988a}:
  \(d = (\bar{X}_A - \bar{X}_B) / s\), where \(\bar{X}_A\),
  \(\bar{X}_B\) are the means of the dependent variable \(X\) for
  samples A and B, while \(s\) is the pooled standard deviation of
  \(X\).  For the case of two continuous variables, the Pearson linear
  correlation coefficient, \(r\), can be used.  In both cases, rules
  of thumb have been developed \citep{Ruscio:2008a} for classifying an
  effect as ``large'' (\(d > 0.8\), \(r > 0.4\)) or ``small''
  (\(d < 0.2\), \(r < 0.1\)).  Alternatively, non-parametric
  statistics can be used, such as the \(A\) measure of stochastic
  superiority \citep{Delaney:2002a}.
\end{enumerate}

Obviously, this depends on the \textit{effect size}, which is the
